---
title: "Portfolio CIS 631"
author: "Michael Bordeaux"
date: "7/26/2023"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidymodels)
```


this is my final project portfolio for CIS 631, in this document I will demonstrate my knowledge of each course objective:



-Describe probability as a foundation of statistical modeling, including inference and maximum likelihood estimation

# Course Objective:

## Determine and apply the appropriate generalized linear model for a specific data context

### showing the course objective:

for this learning objective activity 6 is demonstrating using the appropriate glm for a specfic data context.
```{r include=FALSE}


url <- "https://www.openintro.org/data/csv/resume.csv"

resume <- read_csv(url)


```
exploratory analysis on received_callback variable:
```{r}
ggplot(resume, aes(x = received_callback)) +
  geom_bar() +
  labs(x = "Received Callback", y = "Count")


```


By looking at the above graph we can see that a majority of these resumes did not receive callbacks.

```{r}

resume$received_callback <- factor(resume$received_callback, labels = c("No", "Yes"))

table_data <- table(resume$received_callback)
total <- sum(table_data)
percent <- prop.table(table_data) * 100

table_df <- data.frame(
  received_callback = levels(resume$received_callback),
  n = table_data,
  percent = percent
)

print(table_df)

```

looking at the table above our probablity of a "Yes" is only 8% with an odds .08/(1-.08) of roughly 8% also.

we can further explore this data by adding race into it:

Calculating the probability of a randomly selected person percieved as black it would be ~6% and the odds of a randomly selected resume of a person percieved as black being
called back is .06/(1-.06) roughly also 6%
```{r}
# The {tidymodels} method for logistic regression requires that the response be a factor variable
resume <- resume %>% 
  mutate(received_callback = as.factor(received_callback))

resume_mod <- logistic_reg() %>%
  set_engine("glm") %>%
  fit(received_callback ~ race, data = resume, family = "binomial")

tidy(resume_mod) %>% 
  knitr::kable(digits = 3)
```

regression equation :

y= -2.675 + .438*X + E

to simplify this and look at the equation for corresponding to resumes/persons perceived as black we'd right it as:
y= -2.675 + E

the logg-odds would be: -2.675

and the odds they would be called back is 
roughly .069 or exp(-2.675)

and the probability is .064 of getting called back

linear, trying to fit some sort of a line for some link function 
explore the data and then say the data means this so thats why I chose this model



# Describing the learning objective "Determine and apply the appropriate generalized linear model for a specific data context":

For this learning objective I am using activity 6 to show the determining and applying the appropriate generalized linear model. For this data set it is appropriate to use logistic regression because we have a binary outcome variable. The variable is received_callback which is either a 1 for yes or 0 for no that indicates if a person who submitted there resume received a callback. For the application of this model you can see I first used exploratory graphs to get an understanding of my data set. I think created a table to further analyze my data before approaching this problem. After the initial analysis and more calculating probability and odds I then fit the logistic regression model using glm and passing the binomial attribute. 


-Conduct model selection for a set of candidate models


-Communicate the results of statistical models to a general audience **possibly use the miniproject #1 here?



-Use programming software (i.e., R) to fit and assess statistical models

** will demonstrate with this portfolio project being coded in R



